Combinatorics of the geometry of Wilson loop diagrams II: Grassmann necklaces, dimensions, and denominators

### 2020-11-02 GWLDCII.tex
changes from today's meeting

Thus we have shown that Algorithm \ref{alg:put GN on WLD} does in fact produce a Grassmann necklace. Next we check that it produces the correct one, i.e. the Grassmann necklace that defines $M(W)$. 
\begin{thm}\label{res alg gives correct GN} \todo{fix mixing words and notation (M.28, M.47)} 
The Grassmann necklace $(I_1 \ldots I_n)$ from Theorem~\ref{res:alg gives GN} is the Grassmann necklace of the positroid $M(W)$.
\end{thm}
\begin{proof}
We have shown in the previous theorem that $(I_1,\dots,I_n)$ is a Grassmann necklace; it remains to check that this Grassmann necklace corresponds to the positroid $M(W)$. Recall from the discussion following Definition~\ref{def:grassmann necklace}, and equation \eqref{basesofmatroids} in particular, that it suffices to show the following points:
\begin{itemize}
\item For each $i \in [n]$, $I_i$ is a basis for $M(W)$.
\item If $J$ is lexicographically smaller than $I_i$ with respect to $<_i$ (for any $i \in [n]$) then $J$ is not a basis for $M(W)$. 
\end{itemize}
The first point follows immediately: the algorithm assigns each propagator to a vertex in the support of that propagator, so $I_i$ is a basis for $M(W)$ by Corollary~\ref{lem basis as perm}.

@@

By Corollary~\ref{no coloops} we know that $i \not\in I_1^W$, and so $i$ must label a horizontal edge on the boundary of the Le diagram of $W$, i.e. it must be a column label. The shapes of the Le diagram of $V$ and $W$ are the same except for the insertion of this column, since $I_1$ is the same for $V$ and $W$ except for the incrementation of the indices $\geq i$ in the transition from the necklace for $V$ to the necklace for $W$. 
\end{proof}
\vspace{0.5em}

@@

Also by Lemma~\ref{lem I} we have $I_{n-1}^{W} = I_n^{V} \cup \{n-1\}$. Recall from the proof of Lemma~\ref{lem I} that \hlfix{$n-3,n-2,n-1,n \not\in I_1^V$}{M doesn't like this notation M.58}; therefore
\[
I_1^W\setminus I_{n-1}^W = (I_1^V \setminus I_n^V) \cup\{n-3\},
\]
and $n-3$ is maximal in this set. Similarly, 
\[I_{n-1}^W \setminus I_1^W = (I_n^V \setminus I_1^V) \cup \{n-1\}.
\]
Also $(I_n^V \setminus I_1^V) \cup \{n-1\}\subseteq \{n-1,n\}$ from the definition of Grassmann necklace, and so in particular $n-1$ is minimal in this set. Thus Algorithm~\ref{alg:GN to Le} yields a plus in the $(n-3) \rightarrow (n-1)$ position (i.e. the second box in the bottom row of the Le diagram of $W$) along with any plusses yielded by $I_n^V$.
\end{proof}

@@

We can now describe the plusses contributed by $I_{n-2}^W$ in the remaining two cases: when $n-2 \not\in I_{n-2}^V$, and when $n-2,n-1 \in I_{n-2}^V$.

@@

\item As in Lemma~\ref{lem I} we have $n,n-1,n-2,n-3 \not\in I_1^V$, which implies the first part of the claim. The second part is a consequence of the discussion following equation \eqref{eq necklace}, which shows that $n-4 \not\in I_n^V$.

@@

\item By Lemma~\ref{lem I} we have $I_1^W = I_1^V\cup \{n-3\}$ and $I_{n-2}^W = (I_n^V \setminus \{n-5\}) \cup \{n-1,n-2\}$. By equation \eqref{eq necklace} we have $n-5 \in I_1^V$, and by the following discussion we have $n-1,n-2,n-3 \not\in I_n^V,I_1^V$. It follows that the symmetric difference of $I_1^W$ and $I_{n-2}^W$ is
    \begin{gather*}
    I_1^W \setminus I_{n-2}^W  = (I_1^V\setminus I_n^V) \cup \{n-3, n-5\}. \\
    I_{n-2}^W \setminus I_1^W  = (I_n^V\setminus I_1^V) \cup \{n-1, n-2\}.
\end{gather*}
As per the discussion following equation \eqref{eq necklace} above, we must also have $I_n^V\setminus I_1^V = \{n\}$ and ${I_1^V\setminus I_n^V = \{v\}}$. If $c = n-5$ then the path $v \rightarrow n$ from $I_n^V$ does not intersect with the paths from $\{n-3, n-5\}$ to $\{n-2, n-1\}$ and we obtain the plusses as stated.

@@

Each $r_i$ is the determinant of a minor determined by certian propagators of $W$: specifically, $p$ plays a role in defining $r_i$ if if contributs one set of values (columns) $i^{th}$ element of the Grassmann necklace and a different set to the $(i-1)^{th}$ element. \sanote{major rewrite of this sentence (due to M.62) Karen, look over}. 
%That is, each $r_i$ is the determinant of the minor determined by the propagators (rows) that contribute different values (columns) to the $i^{th}$ element of the Grassmann necklace than they did to the $(i-1)^{th}$. 
Since each $r_i$ is simply the determinant of a particular submatrix of $C(W)$ and an ordering on the propagators has been fixed, the sign of each $r_i$ is well defined.

@@

Next we need to show that $R$ is square-free.  Suppose $f^2|R$.  If $f$ is a linear factor, say from associating propagator $p$ to vertex $a$, then there must be two distinct points in the Grassmann necklace algorithm where $p$ changes from not being associated to vertex $a$ to being associated to vertex $a$.  This contradicts Lemma~\ref{vertex cyclic int lem}.  Now suppose $f$ is a quadratic factor, say from propagators $p$ and $q$ supported on edge $a$ with $p$ clockwise to $q$ on the edge.  In this case it is not possible for any $I_i$ to associate $p$ to $a+1$ and $q$ to $a$.  Furthermore, we know by part (4) that $p$ changes from not being associated to $a$ to being associated to $a$ if and only if $q$ changes from not being associated to $a+1$ to being associated to $a+1$.  Thus $f^2|R$ implies that \hlfix{twice}{Not sure what to do M.72} in the Grassmann necklace $p$ must change from not being associated to vertex $a$ to being associated to vertex $a$. This is again a contradiction, and so $R$ is square-free.

>>>>

Thus we have shown that Algorithm \ref{alg:put GN on WLD} does in fact produce a Grassmann necklace. Next we check that it produces the correct one, i.e. the Grassmann necklace that defines $M(W)$. 
\begin{thm}\label{res alg gives correct GN} 
The Grassmann necklace $(I_1 \ldots I_n)$ from Theorem~\ref{res:alg gives GN} is the Grassmann necklace of the positroid $M(W)$.
\end{thm}
\begin{proof}
We have shown in the previous theorem that $(I_1,\dots,I_n)$ is a Grassmann necklace; it remains to check that this Grassmann necklace corresponds to the positroid $M(W)$. Recall from the discussion following Definition~\ref{def:grassmann necklace}, and equation \eqref{basesofmatroids} in particular, that it suffices to show the following points:
\begin{itemize}
\item For each $i \in [n]$, $I_i$ is a basis for $M(W)$.
\item If $J$ is lexicographically smaller than $I_i$ with respect to $<_i$ (for any $i \in [n]$) then $J$ is not a basis for $M(W)$. 
\end{itemize}
The first point follows immediately: the algorithm assigns each propagator to a vertex in the support of that propagator, so $I_i$ is a basis for $M(W)$ by Corollary~\ref{lem basis as perm}.

@@

By Corollary~\ref{no coloops} we know that $i \not\in I_1^W$, and so $i$ must label a horizontal edge on the boundary of the Le diagram of $W$, i.e. it must be a column label. The shapes of the Le diagram of $V$ and $W$ are the same except for the insertion of this column, since $I_1$ is the same for $V$ and $W$ except for the incrementation of the indices  greater than or equal to $i$ in the transition from the necklace for $V$ to the necklace for $W$.  \todo{fix mixing words and notation (M.28, M.49) throughout, esp wrt <, > $\geq$ etc.} 
\end{proof}
\vspace{0.5em}

@@

Also by Lemma~\ref{lem I} we have $I_{n-1}^{W} = I_n^{V} \cup \{n-1\}$. Recall from the proof of Lemma~\ref{lem I} that none of $n-3,\;n-2,\;n-1$ or $n$ are in $I_1^V$; therefore
\[
I_1^W\setminus I_{n-1}^W = (I_1^V \setminus I_n^V) \cup\{n-3\},
\]
and $n-3$ is maximal in this set. Similarly, 
\[I_{n-1}^W \setminus I_1^W = (I_n^V \setminus I_1^V) \cup \{n-1\}.
\]
Also $(I_n^V \setminus I_1^V) \cup \{n-1\}\subseteq \{n-1,n\}$ from the definition of Grassmann necklace, and so in particular $n-1$ is minimal in this set. Thus Algorithm~\ref{alg:GN to Le} yields a plus in the $(n-3) \rightarrow (n-1)$ position (i.e. the second box in the bottom row of the Le diagram of $W$) along with any plusses yielded by $I_n^V$.
\end{proof}

@@

We can now describe the plusses contributed by $I_{n-2}^W$ in the remaining two cases: when $n-2 \not\in I_{n-2}^V$, and when both $n-2$ and $n-1$ are not in $I_{n-2}^V$.

@@

\item As in Lemma~\ref{lem I} we have that none of $n,\; n-1, \; n-2$ and $n-3$ are in $I_1^V$, which implies the first part of the claim. The second part is a consequence of the discussion following equation \eqref{eq necklace}, which shows that $n-4 \not\in I_n^V$. \todo{propagate M.58 when it appears inline}

@@

\item By Lemma~\ref{lem I} we have $I_1^W = I_1^V\cup \{n-3\}$ and $I_{n-2}^W = (I_n^V \setminus \{n-5\}) \cup \{n-1,n-2\}$. By equation \eqref{eq necklace} we have $n-5 \in I_1^V$, and by the following discussion we have that none of $n-1,\;n-2$ and $n-3$ are not in either $I_n^V$ or $I_1^V$. It follows that the symmetric difference of $I_1^W$ and $I_{n-2}^W$ is
    \begin{gather*}
    I_1^W \setminus I_{n-2}^W  = (I_1^V\setminus I_n^V) \cup \{n-3, n-5\}. \\
    I_{n-2}^W \setminus I_1^W  = (I_n^V\setminus I_1^V) \cup \{n-1, n-2\}.
\end{gather*}
As per the discussion following equation \eqref{eq necklace} above, we must also have $I_n^V\setminus I_1^V = \{n\}$ and ${I_1^V\setminus I_n^V = \{v\}}$. If $c = n-5$ then the path $v \rightarrow n$ from $I_n^V$ does not intersect with the paths from $\{n-3, n-5\}$ to $\{n-2, n-1\}$ and we obtain the plusses as stated.

@@

Each $r_i$ is the determinant of a minor determined by certain propagators of $W$. In particular, if a propagator $p$ contributes one value to the $i^{th}$ element of the Grassmann necklace and a different value to the $(i-1)^{th}$ element, then $p$ is one of the propagators contibuting to the minor. Specifically, the row associated to $p$ and the column associated to the value that $p$ contributes to the $i^{th}$ Grassmann Necklace element appear in the minor.
%That is, each $r_i$ is the determinant of the minor determined by the propagators (rows) that contribute different values (columns) to the $i^{th}$ element of the Grassmann necklace than they did to the $(i-1)^{th}$. 
Since each $r_i$ is simply the determinant of a particular submatrix of $C(W)$ and an ordering on the propagators has been fixed, the sign of each $r_i$ is well defined.

@@

Next we need to show that $R$ is square-free.  Suppose $f^2|R$.  If $f$ is a linear factor, say from associating propagator $p$ to vertex $a$, then there must be two distinct points in the Grassmann necklace algorithm where $p$ changes from not being associated to vertex $a$ to being associated to vertex $a$.  This contradicts Lemma~\ref{vertex cyclic int lem}.  Now suppose $f$ is a quadratic factor, say from propagators $p$ and $q$ supported on edge $a$ with $p$ clockwise to $q$ on the edge.  In this case it is not possible for any $I_i$ to associate $p$ to $a+1$ and $q$ to $a$.  Furthermore, we know by part (4) that $p$ changes from not being associated to $a$ to being associated to $a$ if and only if $q$ changes from not being associated to $a+1$ to being associated to $a+1$.  Thus $f^2|R$ implies that  in the Grassmann necklace $p$ must change twice from not being associated to vertex $a$ to being associated to vertex $a$. This is again a contradiction, and so $R$ is square-free.
